---
title: "for_Olivia_&_Luiz"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
```

Run with caution
```{r}
metadata <- read.csv(file = "../../info/pfam_metadata.csv", header=TRUE, sep=",")

```

```{r}
nested <- metadata %>%
  nest(embedding = c(X0:X1279))
```

```{r}
rm("metadata")
```

Use only the groups that Olivia has sent:
```{r}
chosen_ids <- c(887, 1675, 814, 4493, 5145, 2520, 1791, 1962, 1752, 667, 1721, 1265, 1867, 2340, 1829, 1944, 250, 29, 743, 378)
```

```{r}
filtered <- nested %>%
  filter(Group.ID %in% chosen_ids)

```

```{r}
counts_all <- nested %>%
  group_by(Group.ID) %>%
  count()

counts_filtered <- filtered %>%
  group_by(Group.ID) %>%
  count()
```

```{r}
rm("nested")
```

=====================================================================================
***START HERE******
=====================================================================================


1. We will get just protein name combinations 
```{r}
group <- 29

interest_group_data <- filtered %>%
  filter(Group.ID == group)
```

```{r}
# combinations <- expand.grid(protein1 = interest_group_data$Protein_ID,
#                             protein2 = interest_group_data$Protein_ID)


# here you would also take a subsample (so that soo many combinations are not made)
library(gtools)
interest_group_count <- nrow(interest_group_data)
combinations <- combinations(n = interest_group_count, r = 2, v = interest_group_data$Protein_ID, repeats.allowed = FALSE)
combinations <- as.data.frame(combinations)

combinations <- combinations %>%
  rename("protein_1" = "V1",
         "protein_2" = "V2")

```

Here we get the dataframe with embeddings side by side:
```{r}
protein_1 <- interest_group_data %>%
  select(Protein_ID, embedding) %>%
  rename("protein_1" = Protein_ID)

left_df <- left_join(combinations, protein_1) %>%
  rename("embedding_1" = embedding)

protein_2 <- interest_group_data %>%
  select(Protein_ID, embedding) %>%
  rename("protein_2" = Protein_ID)

all_df <- left_join(left_df, protein_2) %>%
  rename("embedding_2" = embedding) 

```


3. Calculating distance: run with caution (takes forever)
```{r}
CalculateEuclideanDistance <- function(vect1, vect2) sqrt(sum((vect1 - vect2)^2))

with_distance <- all_df %>%
  mutate(distance = map2(all_df$embedding_1, all_df$embedding_2, CalculateEuclideanDistance))
```

```{r}
inner_distances <- with_distance %>%
  select(-c(embedding_1, embedding_2)) %>%
  mutate(group = "within_group")
```

4. We will get new combinations for across groups:
```{r}
# get proteins not including interest group
outer_proteins <- filtered %>%
  filter(!Protein_ID %in% chosen_ids)
```

```{r}
inner_group_count <- nrow(interest_group_data)
outer_protein_sample <- sample_n(outer_proteins, inner_group_count) %>%
  rename("protein_2" = Protein_ID)

combinations2 <- expand.grid(interest_group_data$Protein_ID, outer_protein_sample$protein_2) %>%
  rename("protein_1" = "Var1",
         "protein_2" = "Var2") 

```

joining inner data with outer data:
```{r}
outer_protein_1 <- interest_group_data %>%
  select(Protein_ID, embedding) %>%
  rename("protein_1" = Protein_ID)

outer_left_df <- left_join(combinations2, protein_1) %>%
  rename("embedding_1" = embedding)

outer_protein_2 <- outer_protein_sample %>%
  select(protein_2, embedding)

outer_all_df <- left_join(outer_left_df, outer_protein_2) %>%
  rename("embedding_2" = embedding) 
```

finding distance (Run with caution! takes forever!)
```{r}
with_distance_outer <- outer_all_df %>%
  mutate(distance = map2(outer_all_df$embedding_1, outer_all_df$embedding_2, CalculateEuclideanDistance))
```

```{r}
outer_distances <- with_distance_outer %>%
  select(-c(embedding_1, embedding_2)) %>%
  mutate(group = "across_groups")
```

5. We will combine inner and outer data
```{r}
for_plot <- rbind(inner_distances, outer_distances)
for_plot$distance <- as.numeric(for_plot$distance)
```

6. Making plot:
```{r}
plot <- for_plot %>%
  select(distance, group) %>%
  ggplot(aes(x = distance, fill = group)) +
  geom_density(alpha = 0.5,
               color = NA) +
  #geom_jitter(aes(x = distance, y = 0, height =  0.01)) +
  scale_fill_brewer(palette = "Dark2",
                    labels = c("across groups", "within group")) +
  ggtitle(paste("Group:", group, "Sample size:", inner_group_count)) +
  scale_x_continuous(expand = expansion(add = 0)) +
  scale_y_continuous(expand = expansion(add = 0)) +
  labs(fill = "") +
  theme_cowplot() +
  theme(legend.position = "top")

plot
```

```{r}
ggsave(filename = paste0("../../analysis/figures/for_Olivia_and_Luiz/distances_", group,".png"), plot = plot, width = 7, height = 6)
```
























