---
title: "Touchon_datasets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}

library(tidyverse)
library(cowplot)
library(colorspace)
options(scipen = 999)

```

Loading data:
```{r}
# loading data:
accenssion_phylo_water <- read.csv(file = "../../Touchon_datasets/unfiltered_data/Touchon_accenssions_phylogroups/Cinful_Water_Phylo.csv", header=TRUE, sep=",")
accenssion_phylo_extraintestinal <- read.csv(file = "../../Touchon_datasets/unfiltered_data/Touchon_accenssions_phylogroups/Cinful_Extraintestinal_Phylo.csv", header=TRUE, sep=",")
all_data <- read.csv(file = "../../Touchon_datasets/unfiltered_data/Touchon_accenssions_phylogroups/Australian_Dataset.csv", header=TRUE, sep=",")
CvaB_and_hits <- read.csv(file = "../../Touchon_datasets/unfiltered_data/Supplemental_File_1/Touchon_cinful_output.csv", header=TRUE, sep=",")
hits_only <- read.csv(file = "../../Touchon_datasets/unfiltered_data/Cinful_Touchon_Hmmer_Hits.csv", header=TRUE, sep=",")
accenssions <- read.csv(file = "../../Touchon_datasets/unfiltered_data/1224_Genomes_Cinful.csv", header=TRUE, sep=",")
```

cleaning raw datasets:
```{r}
accenssion_phylo <- all_data %>%
  select(Unique.Name, Strain_Name, Accession.Number, Strain_Category, Phylogroup, Genome_ID) %>%
  rename(accenssion_number = Accession.Number,
         phylogroup = Phylogroup,
         name = Unique.Name,
         strain_category = Strain_Category,
         strain_name = Strain_Name) %>%
  unique()

head(accenssion_phylo)

joined_with_accesssions <- inner_join(accenssion_phylo, accenssions)
```
Calculating microcin hit counts:
```{r}
# cleaning data
microcin_hits <- CvaB_and_hits %>%
  select(sample, component, hmmerHit) %>%
  filter(component == "microcins.verified",
         hmmerHit == "True") %>%
  select(-hmmerHit) %>%
  mutate(strain_name = str_sub(sample, 17, -9))
  
# counting hits per strain/genome
microcin_hit_count <- microcin_hits %>%
  group_by(strain_name) %>%
  mutate(count = n()) %>%
  select(strain_name,  count)

all_data_hits <- CvaB_and_hits %>%
  select(sample) %>%
  mutate(strain_name = str_sub(sample, 17, -9)) %>%
  select(-sample) %>%
  unique()
  
microcin_hit_count_2 <- left_join(all_data_hits, microcin_hit_count) %>%
  mutate(count = ifelse(is.na(count), 0, count)) %>%
  unique() %>%
  rename(microcin_hit_count = count)

```
Calculating CvaB hit counts:
```{r}
# cleaning data
CvaB_hits <- CvaB_and_hits %>%
  select(sample, component, hmmerHit) %>%
  filter(component == "CvaB.verified",
         hmmerHit == "True") %>%
  select(-hmmerHit) %>%
  mutate(strain_name = str_sub(sample, 17, -9))
  
# counting hits per strain/genome
CvaB_hit_count <- CvaB_hits %>%
  group_by(strain_name) %>%
  mutate(count = n()) %>%
  select(strain_name,  count)

all_data_hits <- CvaB_and_hits %>%
  select(sample) %>%
  mutate(strain_name = str_sub(sample, 17, -9)) %>%
  select(-sample) %>%
  unique()
  
CvaB_hit_count_2 <- left_join(all_data_hits, CvaB_hit_count) %>%
  mutate(count = ifelse(is.na(count), 0, count)) %>%
  unique() %>%
  rename(CvaB_hit_count = count)
```

Joining all data into a single dataset:
```{r}
Touchon_data <- inner_join(accenssion_phylo, microcin_hit_count_2)
Touchon_data2 <- inner_join(Touchon_data, CvaB_hit_count_2)

accenssions2 <-accenssions %>%
  select(Genome_ID, Assembly)

Touchon_data3 <- left_join(Touchon_data2, accenssions2) %>%
  distinct(accenssion_number, .keep_all=TRUE)
Touchon_data3

#write.csv(Touchon_data3, "../../Touchon_datasets/filtered_data/cinful_results_for_Touchon_dataset.csv", row.names=FALSE)
```
Getting stats on each group:
```{r}
strain_name_stats <- Touchon_data3 %>% 
  filter(phylogroup == "B2") %>%
  group_by(strain_category) %>%
  summarise(microcin_hits = sum(microcin_hit_count),
            CvaB_hits = sum(CvaB_hit_count)) %>%
  pivot_longer(cols = c(microcin_hits, CvaB_hits), names_to = "cinful_hits", values_to = "hit_count")

#strain_name_stats

strain_plot <- strain_name_stats %>%
  ggplot(aes(y = strain_category, x = hit_count, fill = cinful_hits)) +
  geom_col(position = "dodge", alpha = 0.7) +
  scale_fill_manual(values = c("#bf8943", "#527530")) +
  labs(fill = "") +
  scale_y_discrete(
    name = "",
    expand = expansion(mult = c(0, 0))
  ) +
  scale_x_continuous(
    name = "Cinful hit count",
    expand = expansion(mult = c(0, 0))
  ) +
  theme_minimal_grid()

strain_plot
```

```{r}
phylogroup_stats <- Touchon_data3 %>% 
  group_by(phylogroup) %>%
  summarise(microcin_hits = sum(microcin_hit_count),
            CvaB_hits = sum(CvaB_hit_count)) %>%
  pivot_longer(cols = c(microcin_hits, CvaB_hits), names_to = "cinful_hits", values_to = "hit_count")

#phylogroup_stats

phylo_plot <- phylogroup_stats %>%
  ggplot(aes(y = phylogroup, x = hit_count, fill = cinful_hits)) +
  geom_col(position = "dodge", alpha = 0.7) +
  scale_fill_manual(values = c("#bf8943", "#527530")) +
  labs(fill = "") +
  scale_y_discrete(
    name = "Phylogroup",
    expand = expansion(mult = c(0, 0))
  ) +
  scale_x_continuous(
    name = "Cinful hit count",
    expand = expansion(mult = c(0, 0))
  ) +
  theme_minimal_grid()

phylo_plot
```
Filtering for genomes where a CvaB was found but no microcins were found. 

```{r}
filtered_CvaB <- Touchon_data3 %>%
  filter(CvaB_hit_count == 1,
         microcin_hit_count == 0)
filtered_CvaB

#write.csv(filtered_CvaB, "../../Touchon_datasets/filtered_data/CvaB_found.csv", header=TRUE, sep=",")
```

Compile large dataset (only working with B2):
```{r}
# 4 total entries
only_CvaB_found <- Touchon_data3 %>%
  filter(CvaB_hit_count > 0,
         microcin_hit_count == 0
         #phylogroup == "B2"
         )

#write.csv(only_CvaB_found, "../../Touchon_datasets/filtered_data/Only_CvaB_found.csv", row.names = FALSE)

# none found
only_microcin_found_water <- Touchon_data3 %>%
  filter(CvaB_hit_count == 0,
         microcin_hit_count > 0,
         phylogroup == "B2",
         strain_category == "Water")

# only one found
only_microcin_found_extraintestinal <- Touchon_data3 %>%
  filter(CvaB_hit_count == 0,
         microcin_hit_count > 0,
         phylogroup == "B2",
         strain_category == "Human_Extra-intestinal")

# 5 total entries
both_CvaB_and_microcin_found_water <- Touchon_data3 %>%
  filter(CvaB_hit_count != 0,
         microcin_hit_count != 0,
         phylogroup == "B2",
         strain_category == "Water")

# 33 total entries
nothing_found_water <- Touchon_data3 %>%
  filter(CvaB_hit_count == 0,
         microcin_hit_count == 0,
         phylogroup == "B2",
         strain_category == "Water")

# 35 total entries
both_CvaB_and_microcin_found_extraint <- Touchon_data3 %>%
  filter(CvaB_hit_count != 0,
         microcin_hit_count != 0,
         phylogroup == "B2",
         strain_category == "Human_Extra-intestinal")

# 28 total entries
nothing_found_extraint <- Touchon_data3 %>%
  filter(CvaB_hit_count == 0,
         microcin_hit_count == 0,
         phylogroup == "B2",
         strain_category == "Human_Extra-intestinal")
```

```{r}

for_dataset1 <- nothing_found_extraint %>%
  slice_sample(n = 10) 

for_dataset1_2 <- both_CvaB_and_microcin_found_extraint %>%
  slice_sample(n = 10) 

dataset_1 <- rbind(for_dataset1, for_dataset1_2)

dataset_1_final <- left_join(dataset_1, accenssions2) %>%
  distinct(accenssion_number, .keep_all=TRUE)
dataset_1_final
```

```{r}
nothing_found_water_15 <- nothing_found_water %>%
  slice_sample(n = 15)
  
dataset_water <- rbind(both_CvaB_and_microcin_found_water, nothing_found_water_15)

dataset_water_final <- left_join(dataset_water, accenssions2) %>%
  distinct(accenssion_number, .keep_all=TRUE)
dataset_water_final

```
Idea: 

Dataset 1:
All B2

15 Nothing found (water)--------------------------------------------------cinful: 0, embeddings: 0 (2?)
5 both CvaB and microcin found (water) - there were only 5 of these-------cinful: 8, embeddings: 9 

10 Nothing found (extraintestinal)----------------------------------------cinful: 0, embeddings: 0
10 both CvaB and microcin found (extraintestinal) (use top 10)------------cinful: 10, embeddings: 12 (checking these)

Dataset 2: 
Not all B2

25 only CvaB found and microcin was not found-----------cinful: 0, embeddings: 3 

```{r}
# plot for only CvaB found:

phylogroup_stats <- Touchon_data2 %>% 
  filter(CvaB_hit_count > 0,
         microcin_hit_count == 0) %>%
  group_by(phylogroup) %>%
  summarise(CvaB_hits = sum(CvaB_hit_count)) 
#phylogroup_stats

phylo_plot <- phylogroup_stats %>%
  ggplot(aes(y = phylogroup, x = CvaB_hits)) +
  geom_col(position = "dodge", 
           alpha = 0.7,
           fill = "#bf8943") +
  scale_y_discrete(
    name = "Phylogroup",
    expand = expansion(mult = c(0, 0))
  ) +
  scale_x_continuous(
    name = "Cinful hit count",
    
    expand = expansion(mult = c(0, 0))
  ) +
  geom_text(position = position_stack(vjust = .9), 
            aes(label = CvaB_hits)) +
  theme_cowplot()

phylo_plot

```
```{r}
strain_name_stats <- Touchon_data2 %>% 
  filter(CvaB_hit_count > 0,
         microcin_hit_count == 0) %>%
  group_by(strain_category) %>%
  summarise(CvaB_hits = sum(CvaB_hit_count)) 

#strain_name_stats

strain_plot <- strain_name_stats %>%
  ggplot(aes(y = strain_category, x = CvaB_hits)) +
 geom_col(position = "dodge", 
           alpha = 0.7,
           fill = "#bf8943") +
  scale_y_discrete(
    name = "",
    expand = expansion(mult = c(0, 0))
  ) +
  scale_x_continuous(
    name = "Cinful hit count",
    
    expand = expansion(mult = c(0, 0))
  ) +
  geom_text(position = position_stack(vjust = .9), 
            aes(label = CvaB_hits)) +
  theme_cowplot()

strain_plot
```

